Exploring Homicide Trends in Toronto: A Seventeen-Year Statistical Overview (2004-2020)

### 2024-01-20 outputs/paper/paper.qmd
Grammar fixes.

In addressing this need, the study utilizes a robust dataset from OpenDataToronto [@opendatatoronto], specifically the “Police Annual Statistical Report - Homicides,” published by the Toronto Police Services. This dataset provides a detailed account of all recorded homicides under the jurisdiction of the Toronto Police Services over a seventeen-year period. Our research offers a meticulous examination of both the spatial distribution of homicides across Toronto's neighbourhoods and the methods used in these violent acts. The analysis, conducted through the R programming language [@R] and supplemented with various analytical tools like Tidyverse [@tidyverse], Janitor [@janitor], and Knitr [@knitr], delves deep into the trends, revealing patterns that could have implications for policy formulation and implementation.

@@

The dataset utilized in this paper is the aforementioned “Police Annual Statistical Report - Homicides”, which includes data from every homicide under the Toronto Police Services jurisdiction between 2004 and 2020. The data set itself contains 1166 samples each with 10 variables. Because Toronto Police Services is the only institution qualified to deal with homicides, no other datasets were considered. One potentially significant concern with this data ,is the possibility that it is incomplete due to undetected homicides. However, I do not believe that there are sufficient undetected homicides to greatly impact the ensuing analysis. Another problem with the data lies in how it handles the location. The Toronto Police Services have deliberately made the location of each occurrence at the nearest intersection, and therefore the data is not flawless by the admission of the data procurers. Finally, the relatively small number of homicides per year in Toronto renders the data somewhat difficult to analyze. When trying to examine trends within certain communities and regions, the deviation was too high to extrapolate anything of interest.

@@

The data includes the date of the event, the type of homicide, as well as the neighbourhood in which the homicide took place. The date is unambiguous and needs no further explanation. The type of homicide is broken into three clear sub-categories which are “shooting”, “stabbing” and “other”. For neighbourhood, Toronto is broken up into 158 “social planning neighbourhoods” [@neighbourhoods]. These are each numbered and reported independently for the purposes of this paper, (this is also how StatsCan breaks Toronto into neighbourhoods for all of their relevant data sets).

With regard to measurement the precision of the location is inaccurate, but the important aspects of the data set are unambiguous in terms of their ability to be measured. A homicide is a countable event that does not need to be measured, the categorization of homicide is the same.  Only the date may have measurement errors but these would be unimpactful for the following analysis.

@@

Our analysis delves into the details of homicide trends in the Greater Toronto Area (GTA) from 2004 to 2020, revealing insights about the nature and distribution of these violent incidents.

>>>>

In addressing this need, the study utilizes a robust dataset from OpenDataToronto [@opendatatoronto], specifically the “Police Annual Statistical Report - Homicides,” published by the Toronto Police Services. This dataset provides a detailed account of all recorded homicides under the jurisdiction of the Toronto Police Services over a seventeen-year period. Our research offers an examination of both the spatial distribution of homicides across Toronto's neighbourhoods and the methods used in these violent acts. The analysis, conducted through the R programming language [@R] and supplemented with various analytical tools like Tidyverse [@tidyverse], Janitor [@janitor], and Knitr [@knitr], delves deep into the trends, revealing patterns that could have implications for policy formulation and implementation.

@@

The dataset utilized in this paper is the aforementioned “Police Annual Statistical Report - Homicides”, which includes data from every homicide under the Toronto Police Services jurisdiction between 2004 and 2020. The data set itself contains 1166 samples each with 10 variables. Because Toronto Police Services is the only institution qualified to deal with homicides, no other datasets were considered. One potentially significant concern with this data, is the possibility that it is incomplete due to undetected homicides. However, I do not believe that there are sufficient undetected homicides to greatly impact the ensuing analysis. Another problem with the data lies in how it handles the location. The Toronto Police Services have deliberately made the location of each occurrence at the nearest intersection, and therefore the data is not flawless by the admission of the data procurers. Finally, the relatively small number of homicides per year in Toronto renders the data somewhat difficult to analyze. When trying to examine trends within certain communities and regions, the deviation was too high to extrapolate anything of interest.

@@

The data includes the date of the event, the type of homicide, as well as the neighbourhood in which the homicide took place. The date is unambiguous and needs no further explanation. The type of homicide is broken into three clear sub-categories which are “shooting”, “stabbing” and “other”. For neighbourhood, Toronto is broken up into 158 “social planning neighbourhoods” [@neighbourhoods]. These are each numbered and reported independently for the purposes of this paper, (this is also how Statistics Canada breaks Toronto into neighbourhoods for all of their relevant data sets).

With regard to measurement, the precise location is inaccurate, but the important aspects of the data set are unambiguous in terms of their ability to be measured. A homicide is a countable event that does not need to be measured, the categorization of homicide is the same.  Only the date may have measurement errors but these would be unimpactful for the following analysis.

@@

Our analysis delves into the details of homicide trends in the GTA from 2004 to 2020, revealing insights about the nature and distribution of these violent incidents.
